---
title: "Shiny practical"
author: "Rhian Davies"
date: "04/03/2020"
output: html_document
---


## The data

We're going to work with the IMDB movies data. 
It lives in the ggplotmovies package. You can install the package by running:

```{r, eval = FALSE}
install.packages("ggplot2movies")
```

To load the dataset into your R session, run the following command

```{r, eval = FALSE}
data("movies", package = "ggplot2movies")
```

You should now have an object called `movies` in your environment. 
Have a look at it.

For our first shiny app, we are going to make a simple shiny app which allows the user to choose what colour the points in a scatter plot are going to be.

## Input

The input will be a simple radio button, with a three colours to choose from. 

Look at the [widget gallery](https://shiny.rstudio.com/gallery/widget-gallery.html), which widget do you need?

Look at the code on the widget page, and adapt it to create a drop down menu with three different colour options. To see a full list of all of R's inbuilt colours, run `colors()`.

```{r, eval = FALSE}
  radioButtons("colour", label = h3("Choose a colour"),
    choices = list("Purple" = "purple", "Fire Brick" = "firebrick2", "Maroon" = "maroon4") ,
    selected = 1)
```

## Output 

For our output plot we'll create a plot showing the relationship between movie length and movie budget. 

Create a scatter plot in ggplot2 with length on the x axis and rating on the y axis. 

There is one movie which is **ridiculously** long, so I suggest you run `filter(length < 300)` before you plot to exclude that outlier.

```{r, eval = FALSE}
movies %>%
  filter(length < 300) %>%
  ggplot(aes(x = length, y = rating)) + 
  geom_point(col = "purple")
```

## Putting it all together

Now we're going to put it all together. Copy the code below, and paste it into a file called `app.R`.

```{rapp, eval = FALSE}
library(shiny)
library(tidyverse)

ui <- fluidPage()

server <- function(input, output) {}

shinyApp(ui = ui, server = server)
```

### The server

Let's start with the server and recap the 3 rules for working the server.

* Save objects to display to output$
* Build objects to display with render*()
* Access input values with input$

Your server function should therefore look something like this.

```{r server, eval = FALSE}
server <- function(input, output) {
  library(ggplot2)
  
  #Load the data
  data("movies", package = "ggplot2movies")
  
  # Create the scatter and save to output$
  output$scatterplot = renderPlot({
    movies %>%
      filter(length < 300) %>%
      ggplot(aes(x = length, y = rating)) + 
      geom_point(col = input$colour)
  })
}
```

### The UI

* Copy the code for the radio buttons into the `fluidPage()`
* Copy the code to show he output plot into `fluidPage()` and separate the two lines with a comma. 
* Remember all we need is `plotOutput("scatterplot")`

Your code for your UI should look something like this:

```{r UI, eval = FALSE}
ui <- fluidPage(radioButtons("colour", label = h3("Choose a colour"),
                             choices = list("Purple" = "purple",
                                            "Fire Brick" = "firebrick2",
                                            "Maroon" = "maroon4") ,
                             selected = "purple"),
                plotOutput("scatterplot")
)
```

Run your shiny app. Remember you need to press the STOP button if you want to start using R again. 

## Further questions

1. Add a text box input which allows the user type in a title for the plot. Notice that the title updates letter-by-letter. Why is that?

1. Add an input select box which allows the user to select a particular mpaa classification for the movies (R, NC-17, PG, PG-13). When the user selects a classification, the ggplot should show only the movies with that classification.

1. Add a slider range input that allows the user to select the years that the films should be shown from.